---
title: "PowerAnalysisSimulation"
author: "Antonia Krefeld-Schwalb"
date: "26 juin 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Input values:
True Effect sizes : | Hypothesized Effect Size :
--- | --- | ---
True Effect sizes : |  unbiased | biased
--- | --- | ---
 $\hat\delta= .2$| $\delta = .2$|$\delta  = .5$


*based on the review of published AND unpublished studies as well as theoretical considerations
**publication bias


```{r }

d_true <- c(.2, .5, 1, 10)

d_hyp <- c(.2, .5, NA)
```


#Sample Size

Set error rates as coventional to $\alpha = .05$ and $\beta = .20$ and calculate effect size based on power analysis. If no effect sizes has been hypothesized, N is randomly chosen between 0 and 100. 


#Simulate t-value

```{r }

alpha = .05
beta = .20
pow = 1-beta
Nsamples = 1000

N = rep(NA, length(d_hyp)) # Sample size needed to achieve power
p = Tval = TvalLow = d_emp = d_empLow = L0x = LR= Lhypx =Lhyp = L1x = array(NA, dim = c(length(d_hyp), Nsamples, 2, length(d_true))) 

for (k in 1:length(d_true)){

  for (dd in 1:length(d_hyp)){
    
      # N[dd] = round(((qnorm(1-alpha)+qnorm(pow))/d_hyp[dd]*sqrt(2))^2 ) }
  
      for (ss in 1:Nsamples){ # for variation
        
        # if d_hyp IS NOT NA => estimate sample size based on alpha, beta to find effect (to reject H0: effect is not       significant) => min sample size needed to reject H0 at 0.05
        
      if (!is.na(d_hyp[dd])){
        # N[dd] = round(((qnorm(1-alpha)+qnorm(pow))/(d_hyp[dd])*(sqrt(2)))^2) }
        N[dd] = round(2*(qnorm(1-alpha/2)+qnorm(pow))^2/d_hyp[dd]^2)
        }
  
      # if d_hyp IS NA => if no d_hyp => take random sample size (no hyp in marketing research) to simulate usual way of       consumer researcher
      if (is.na(d_hyp[dd])){
        N[dd] = sample(seq(10,100),1)
        }
      #mean(p[1,,1]<.05)
  
      for (times in 1:2){
        x1 = rnorm(N[dd], d_true[k], 1) # expectation = d_true, sd = 1, N samples
        meanx1 = mean(x1)
  
        #Distribution of Control condition  
        x0 = rnorm(N[dd], 0, 1) # expectation = 0, sd = 1, N samples
        meanx0 = mean(x0)
      
        #calculate Standard Error as in R t.test function
        v = ((N[dd]-1)*var(x1)+(N[dd]-1)*var(x0))/(2*N[dd]-2) #not SE but avg deviation between 2 groups (pooled var)
      
        SE =  sqrt(v*(1/N[dd]+1/N[dd]))
        lowdiff = mean(x1 - x0) - 1.96*SE #different way of estimating effect size: more conservative (lower bound of 95% CI) => IGNORE
  
        #Calculate t-values
        Tval[dd,ss,times, k] = (meanx1-meanx0)/(sqrt((var(x1) + var(x0))/2)*sqrt(2/N[dd]))
        p[dd,ss,times, k] = 2*pt(-Tval[dd,ss,times, k], (2*N[dd]-2)) # p-value
        d_emp[dd,ss,times, k] = 2*Tval[dd,ss,times, k]/sqrt(N[dd]) #d-emp
      
        TvalLow[dd,ss,times, k] = (lowdiff)/(sqrt((var(x1) + var(x0))/2)* sqrt(2/N[dd])) # lower bounds (effect based on lower bound) => IGNORE
        d_empLow[dd,ss,times, k] = 2*TvalLow[dd,ss,times, k]/sqrt(N[dd]) # lower bounds => IGNORE
  
        L0x[dd,ss,times, k] = dt(Tval[dd,ss,times, k], (2*N[dd]-2), 0) # likelihood of 0 given data, density at observed t-value given df and non centrality param (0)
      
        if(!is.na(d_hyp[dd])){
          
          ncp = d_hyp[dd]*sqrt(N[dd])/sqrt(2) #non-centrality param (of t distribution)
          L1x[dd,ss,times, k] = dt(Tval[dd,ss,times, k], (2*N[dd]-2), ncp) #likelihood of obs t-value given hypothesied effect           size for transformed distr given ncp
          Lhypx[dd,ss,times, k] = dt(ncp, (2*N[dd]-2), ncp) #density at median of same distribution (.39)
          LR[dd,ss,times, k] = L1x[dd,ss,times, k]/L0x[dd,ss,times, k] # likelihood ratio (obs t-value and null hyp) (LRT>4 => obs effect size is not from the H0 distr)
          Lhyp[dd,ss,times, k] = Lhypx[dd,ss,times, k]/L1x[dd,ss,times, k] #likelihood ration of the median of hypothesied distr to obs effect size => how far obs effect size from the H0 effect size
        }
      }
    }
  }
}
``` 

Average ratio of effect sizes if $\it{p_1} < .05 :\frac{d_{emp_2}}{d_{emp_1}}$:

$d_{hyp} =$ `r d_hyp[1]` \& $d_{true} =$ `r d_true` 

`r mean(d_emp[1,which(p[1,,1] <.05),2]/d_emp[1,which(p[1,,1] <.05),1])`


$d_{hyp} =$ `r d_hyp[2]` \& $d_{true} =$ `r d_true` 

`r mean(d_emp[2,which(p[2,,1] <.05),2]/d_emp[2,which(p[2,,1] <.05),1])`


$d_{hyp} =$ `r d_hyp[3]` \& $d_{true} =$ `r d_true` 

`r mean(d_emp[3,which(p[3,,1] <.05),2]/d_emp[3,which(p[3,,1] <.05),1])`



Average ratio of effect sizes if $LR > \frac{\beta}{\alpha} =$ 'r beta/alpha' : $\frac{d_{emp_2}}{d_{emp_1}}$:

$d_{hyp} =$ `r d_hyp[1]` \& $d_{true} =$ `r d_true` 

`r mean(d_emp[1,which(LR[1,,1] >20),2]/d_emp[1,which(LR[1,,1] >20),1])`


$d_{hyp} =$ `r d_hyp[2]` \& $d_{true} =$ `r d_true` 

`r mean(d_emp[2,which(LR[1,,1] >20),2]/d_emp[2,which(LR[1,,1] >20),1])`




Average ratio of effect sizes if $LR > \frac{\beta}{\alpha} =$ 'r beta/alpha' : $\frac{d_{emp_2}}{d_{emp_1}}$:

$d_{hyp} =$ `r d_hyp[1]` \& $d_{true} =$ `r d_true` 

`r mean(d_emp[1,which(LR[1,,1] >20 & Lhyp[1,,1] > 4 ),2]/d_hyp[1])`


$d_{hyp} =$ `r d_hyp[2]` \& $d_{true} =$ `r d_true` 

`r mean(d_emp[1,which(LR[2,,1] >20 & Lhyp[2,,1] > 4 ),2]/d_hyp[2])`

