---
title: "PowerAnalysisSimulation"
author: "Antonia Krefeld-Schwalb"
date: "26 juin 2019"
output: html_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Input values:
True Effect sizes : | Hypothesized Effect Size :
  --- | --- | ---
  True Effect sizes : |  unbiased | biased
--- | --- | ---
  $\hat\delta= .2$| $\delta = .2$|$\delta  = .5$
  
  
  *based on the review of published AND unpublished studies as well as theoretical considerations
**publication bias
```{r }
d_true <- c(0, .2, .5, .8)
d_hyp <- c(.2, .5, .8, NA)
```

#Sample Size

Set error rates as coventional to $\alpha = .05$ and $\beta = .20$ and calculate effect size based on power analysis. If no effect sizes has been hypothesized, N is randomly chosen between 0 and 100. 


#Simulate t-value
```{r }
alpha = .05
beta = .20
pow = 1-beta
Nsamples = 1000
N = rep(NA, length(d_hyp)) # Sample size needed to achieve power
p = Tval = TvalLow = d_emp = d_empLow = L0x = LR= Lhypx =Lhyp = L1x = array(NA, dim = c(length(d_hyp), Nsamples, 2, length(d_true))) 

df1 =df2 =matrix(0,nrow = 1, ncol=5)

```

Simulate `r Nsamples` t-test, twice. The first t-test represents the scientific study and the second t-test can be regarded as the pratical implementation.

```{r }

  for (dd in 1:length(d_hyp)){
    for (k in 1:length(d_true)){
  # repeat the Simulation Nsamples times
    for (ss in 1: Nsamples){
      # IF existing point Hypothesis, estimate minimum sample size for errorates alpha and beta
      if (!is.na(d_hyp[dd])){
          N[dd] = round(2*(qnorm(1-alpha/2)+qnorm(pow))^2/d_hyp[dd]^2) }
      # IF not existing point Hypothesis, draw random sample size from sequence 10 - 100
      if (is.na(d_hyp[dd])){
          N[dd] = sample(seq(10,100),1)}

      
# repeat the simulation twice: 1.lab 2. pratical implementation
for (times in 1:2){
  # simulate distribution oftreatment group
  x1 =rnorm(N[dd],d_true[k], 1)
  meanx1 = mean(x1)
  # simulate distribution oftreatment group
  x0 =rnorm(N[dd],0, 1)
  meanx0 = mean(x0)
      

        #Calculate t-values
        Tval[dd,ss,times, k] = (meanx1-meanx0)/(sqrt((var(x1) + var(x0))/2)*sqrt(2/N[dd]))
        p[dd,ss,times, k] = 2*pt(-Tval[dd,ss,times, k], (2*N[dd]-2)) # p-value
        d_emp[dd,ss,times, k] = 2*Tval[dd,ss,times, k]/sqrt(N[dd]) #d-emp
        if(times == 1) df1 = rbind(df1,c(d_emp[dd,ss,times, k],  p[dd,ss,times, k], d_hyp[dd],d_true[k],times))
        if(times == 2) df2 = rbind(df2,c(d_emp[dd,ss,times, k],  p[dd,ss,times, k], d_hyp[dd],d_true[k],times))
        
###### the following lines of code are only relevant later for LR test and test on the lower borderof the interval
#   
#   #calculate Standard Error as in R t.test function
        # TvalLow[dd,ss,times, k] = (lowdiff)/(sqrt((var(x1) + var(x0))/2)* sqrt(2/N[dd])) # lower bounds (effect based on lower bound) => IGNORE
        # d_empLow[dd,ss,times, k] = 2*TvalLow[dd,ss,times, k]/sqrt(N[dd]) # lower bounds => IGNORE
        # 
        # L0x[dd,ss,times, k] = dt(Tval[dd,ss,times, k], (2*N[dd]-2), 0) # likelihood of 0 given data, density at observed t-value given df and non centrality param (0)
        # 
        # if(!is.na(d_hyp[dd])){
        # 
        #   ncp = d_hyp[dd]*sqrt(N[dd])/sqrt(2) #non-centrality param (of t distribution)
        #   L1x[dd,ss,times, k] = dt(Tval[dd,ss,times, k], (2*N[dd]-2), ncp) #likelihood of obs t-value given hypothesied effect           size for transformed distr given ncp
        #   Lhypx[dd,ss,times, k] = dt(ncp, (2*N[dd]-2), ncp) #density at median of same distribution (.39)
        #   LR[dd,ss,times, k] = L1x[dd,ss,times, k]/L0x[dd,ss,times, k] # likelihood ratio (obs t-value and null hyp) (LRT>4 => obs effect size is not from the H0 distr)
        #   Lhyp[dd,ss,times, k] = Lhypx[dd,ss,times, k]/L1x[dd,ss,times, k] #likelihood ration of the median of hypothesied distr to obs effect size => how far obs effect size from the H0 effect size
        # }
      }
    }
  }
}
``` 


The plot illustrates the observed effect sizes in the two test, if the first test was significant. The column indicates the true effect size and the row indicates the hypothesized effect size, which was used to determine the sample size of the test, to achieve $\beta = .8$ with $\alpha = .05$. The color indicates the ratio of the two effect sizes and the diagonal goes through zero and one. 

```{r }
library(ggplot2)
library(knitr)


df1 = df1[-1,]
df2 = df2[-1,]

df= cbind(df1[which(df1[,2] < .05), c(1,3,4)],df2[which(df1[,2] < .05), 1])



colnames(df)= c("d_1","d_hyp","d_true","d_2")
df= as.data.frame(df)

df$ratio = df$d_1/df$d_2
df$ratio[df$d_1 ==0] = df$df_2/.0001
df$ratioFactor[df$ratio < .5]=1
df$ratioFactor[df$ratio >= .5]=2
df$ratioFactor[df$ratio >= 1] =3
df$ratioFactor[df$ratio >= 1.5]=4

summaries = matrix(NA, nrow= length(d_true)*length(d_hyp), ncol = 4)

for(k in 1:length(d_true)){
  for (dd in 1:length(d_hyp)){
    print(length(d_hyp)*(k-1)+dd)
    summaries[(length(d_hyp)*(k-1) + dd) ,] = c(round(mean(d_emp[dd,which(p[dd,,1,k] <.05),2,k]/d_emp[dd,which(p[dd,,1,k] <.05),1,k]),2),
                                                round(sd(d_emp[dd,which(p[dd,,1,k] <.05),2,k]/d_emp[dd,which(p[dd,,1,k] <.05),1,k]),2), d_hyp[dd],d_true[k])
    
  }}
summaries = as.data.frame(summaries)
colnames(summaries)= c(expression(M(d_[emp_2]/d_[emp_1])),expression(SD(d_[emp_2]/d_[emp_1])),"d_hyp","d_true")


ggplot(df, aes(x=d_1, y=d_2, color = factor(ratioFactor)))+
  geom_point()+
  facet_grid(d_hyp~d_true)+
  xlab(bquote(d_[obs1]))+
  ylab(bquote(d_[obs2]))+
  geom_abline(intercept = 0, slope = 1)+
  scale_color_discrete(name = "", labels = c(bquote(d_[emp_2]/d_[emp_1]<.5),
                                                                       bquote(d_[emp_2]/d_[emp_1]<1),
                                                                       bquote(d_[emp_2]/d_[emp_1]<1.5),
 
                                                                                                                   bquote(d_[emp_2]/d_[emp_1]>=1.5)))
```


Average ratio of effect sizes in study 1 and study 2 $\it{p_1} < .05$:


```{r }

kable(summaries,col.names = c("M(ratio)",
                              "SD(ratio)",
                              "d_hyp","d_true"))


# 
# $d_{hyp} =$ `r d_hyp[1]` \& $d_{true} =$ `r d_true[1]` 
# 
# $M(\frac{d_{emp_2}}{d_{emp_1}})$= `r round(mean(d_emp[1,which(p[1,,1,1] <.05),2,1]/d_emp[1,which(p[1,,1,1] <.05),1,1]),2)`, $SD(\frac{d_{emp_2}}{d_{emp_1}})$= `r round(sd(d_emp[1,which(p[1,,1,1] <.05),2,1]/d_emp[1,which(p[1,,1,1] <.05),1,1]),2)`
# 
# $d_{hyp} =$ `r d_hyp[2]` \& $d_{true} =$ `r d_true[1]` 
# 
# $M(\frac{d_{emp_2}}{d_{emp_1}})$= `r round(mean(d_emp[2,which(p[2,,1,1] <.05),2,1]/d_emp[2,which(p[2,,1,1] <.05),1,1]),2)`, $SD(\frac{d_{emp_2}}{d_{emp_1}})$= `r round(sd(d_emp[2,which(p[2,,1,1] <.05),2,1]/d_emp[2,which(p[2,,1,1] <.05),1,1]),2)`
# 
# 
# $d_{hyp} =$ `r d_hyp[3]` \& $d_{true} =$ `r d_true[1]` 
# 
# $M(\frac{d_{emp_2}}{d_{emp_1}})$= `r round(mean(d_emp[3,which(p[3,,1,1] <.05),2,1]/d_emp[3,which(p[3,,1,1] <.05),1,1]),2)`, $SD(\frac{d_{emp_2}}{d_{emp_1}})$= `r round(sd(d_emp[3,which(p[3,,1,1] <.05),2,1]/d_emp[3,which(p[3,,1,1] <.05),1,1]),2)`
# 
# $d_{hyp} =$ `r d_hyp[4]` \& $d_{true} =$ `r d_true[1]` 
# 
# $M(\frac{d_{emp_2}}{d_{emp_1}})$= `r round(mean(d_emp[4,which(p[4,,1,1] <.05),2,1]/d_emp[4,which(p[4,,1,1] <.05),1,1]),2)`, $SD(\frac{d_{emp_2}}{d_{emp_1}})$= `r round(sd(d_emp[4,which(p[4,,1,1] <.05),2,1]/d_emp[4,which(p[4,,1,1] <.05),1,1]),2)`
# 
# 
# 
# 
# $d_{hyp} =$ `r d_hyp[1]` \& $d_{true} =$ `r d_true[2]` 
# 
# $M(\frac{d_{emp_2}}{d_{emp_1}})$= `r round(mean(d_emp[1,which(p[1,,1,2] <.05),2,2]/d_emp[1,which(p[1,,1,2] <.05),1,2]),2)`, $SD(\frac{d_{emp_2}}{d_{emp_1}})$= `r round(sd(d_emp[1,which(p[1,,1,2] <.05),2,2]/d_emp[1,which(p[1,,1,2] <.05),1,2]),2)`
# 
# $d_{hyp} =$ `r d_hyp[2]` \& $d_{true} =$ `r d_true[2]` 
# 
# $M(\frac{d_{emp_2}}{d_{emp_1}})$= `r round(mean(d_emp[2,which(p[2,,1,2] <.05),2,2]/d_emp[2,which(p[2,,1,2] <.05),1,2]),2)`, $SD(\frac{d_{emp_2}}{d_{emp_1}})$= `r round(sd(d_emp[2,which(p[2,,1,2] <.05),2,2]/d_emp[2,which(p[2,,1,2] <.05),1,2]),2)`
# 
# 
# $d_{hyp} =$ `r d_hyp[3]` \& $d_{true} =$ `r d_true[2]` 
# 
# $M(\frac{d_{emp_2}}{d_{emp_1}})$= `r round(mean(d_emp[3,which(p[3,,1,2] <.05),2,2]/d_emp[3,which(p[3,,1,2] <.05),1,2]),2)`, $SD(\frac{d_{emp_2}}{d_{emp_1}})$= `r round(sd(d_emp[3,which(p[3,,1,2] <.05),2,2]/d_emp[3,which(p[3,,1,2] <.05),1,2]),2)`
# 
# 
# $d_{hyp} =$ `r d_hyp[4]` \& $d_{true} =$ `r d_true[2]` 
# 
# $M(\frac{d_{emp_2}}{d_{emp_1}})$= `r round(mean(d_emp[4,which(p[4,,1,2] <.05),2,2]/d_emp[4,which(p[4,,1,2] <.05),1,2]),2)`, $SD(\frac{d_{emp_2}}{d_{emp_1}})$= `r round(sd(d_emp[4,which(p[4,,1,2] <.05),2,2]/d_emp[4,which(p[4,,1,2] <.05),1,2]),2)`
# 
# 
# 
# $d_{hyp} =$ `r d_hyp[1]` \& $d_{true} =$ `r d_true[3]` 
# 
# $M(\frac{d_{emp_2}}{d_{emp_1}})$= `r round(mean(d_emp[1,which(p[1,,1,3] <.05),2,3]/d_emp[1,which(p[1,,1,3] <.05),1,3]),2)`, $SD(\frac{d_{emp_2}}{d_{emp_1}})$= `r round(sd(d_emp[1,which(p[1,,1,3] <.05),2,3]/d_emp[1,which(p[1,,1,3] <.05),1,3]),2)`
# 
# $d_{hyp} =$ `r d_hyp[2]` \& $d_{true} =$ `r d_true[3]` 
# 
# $M(\frac{d_{emp_2}}{d_{emp_1}})$= `r round(mean(d_emp[2,which(p[2,,1,3] <.05),2,3]/d_emp[2,which(p[2,,1,3] <.05),1,3]),2)`, $SD(\frac{d_{emp_2}}{d_{emp_1}})$= `r round(sd(d_emp[2,which(p[2,,1,3] <.05),2,3]/d_emp[2,which(p[2,,1,3] <.05),1,3]),2)`
# 
# 
# $d_{hyp} =$ `r d_hyp[3]` \& $d_{true} =$ `r d_true[3]` 
# 
# $M(\frac{d_{emp_2}}{d_{emp_1}})$= `r round(mean(d_emp[3,which(p[3,,1,3] <.05),2,3]/d_emp[3,which(p[3,,1,3] <.05),1,3]),2)`, $SD(\frac{d_{emp_2}}{d_{emp_1}})$= `r round(sd(d_emp[3,which(p[3,,1,3] <.05),2,3]/d_emp[3,which(p[3,,1,3] <.05),1,3]),2)`
# 
# $d_{hyp} =$ `r d_hyp[4]` \& $d_{true} =$ `r d_true[3]` 
# 
# $M(\frac{d_{emp_2}}{d_{emp_1}})$= `r round(mean(d_emp[4,which(p[4,,1,3] <.05),2,3]/d_emp[4,which(p[4,,1,3] <.05),1,3]),2)`, $SD(\frac{d_{emp_2}}{d_{emp_1}})$= `r round(sd(d_emp[4,which(p[4,,1,3] <.05),2,3]/d_emp[4,which(p[4,,1,3] <.05),1,3]),2)`
# 
# 
# $d_{hyp} =$ `r d_hyp[1]` \& $d_{true} =$ `r d_true[4]` 
# 
# $M(\frac{d_{emp_2}}{d_{emp_1}})$= `r round(mean(d_emp[1,which(p[1,,1,4] <.05),2,4]/d_emp[1,which(p[1,,1,4] <.05),1,4]),2)`, $SD(\frac{d_{emp_2}}{d_{emp_1}})$= `r round(sd(d_emp[1,which(p[1,,1,4] <.05),2,4]/d_emp[1,which(p[1,,1,4] <.05),1,4]),2)`
# 
# $d_{hyp} =$ `r d_hyp[2]` \& $d_{true} =$ `r d_true[4]` 
# 
# $M(\frac{d_{emp_2}}{d_{emp_1}})$= `r round(mean(d_emp[2,which(p[2,,1,4] <.05),2,4]/d_emp[2,which(p[2,,1,4] <.05),1,4]),2)`, $SD(\frac{d_{emp_2}}{d_{emp_1}})$= `r round(sd(d_emp[2,which(p[2,,1,4] <.05),2,4]/d_emp[2,which(p[2,,1,4] <.05),1,4]),2)`
# 
# 
# $d_{hyp} =$ `r d_hyp[3]` \& $d_{true} =$ `r d_true[4]` 
# 
# $M(\frac{d_{emp_2}}{d_{emp_1}})$= `r round(mean(d_emp[3,which(p[3,,1,4] <.05),2,4]/d_emp[3,which(p[3,,1,4] <.05),1,4]),2)`, $SD(\frac{d_{emp_2}}{d_{emp_1}})$= `r round(sd(d_emp[3,which(p[3,,1,4] <.05),2,4]/d_emp[3,which(p[3,,1,4] <.05),1,4]),2)`
# 
# $d_{hyp} =$ `r d_hyp[4]` \& $d_{true} =$ `r d_true[4]` 
# 
# $M(\frac{d_{emp_2}}{d_{emp_1}})$= `r round(mean(d_emp[4,which(p[4,,1,4] <.05),2,4]/d_emp[4,which(p[4,,1,4] <.05),1,4]),2)`, $SD(\frac{d_{emp_2}}{d_{emp_1}})$= `r round(sd(d_emp[4,which(p[4,,1,4] <.05),2,4]/d_emp[4,which(p[4,,1,4] <.05),1,4]),2)`
# Average ratio of effect sizes if $LR > \frac{\beta}{\alpha} =$ 'r beta/alpha' : $\frac{d_{emp_2}}{d_{emp_1}}$:
#   $d_{hyp} =$ `r d_hyp[1]` \& $d_{true} =$ `r d_true` 
# `r mean(d_emp[1,which(LR[1,,1] >20),2]/d_emp[1,which(LR[1,,1] >20),1])`
# $d_{hyp} =$ `r d_hyp[2]` \& $d_{true} =$ `r d_true` 
# `r mean(d_emp[2,which(LR[1,,1] >20),2]/d_emp[2,which(LR[1,,1] >20),1])`
# Average ratio of effect sizes if $LR > \frac{\beta}{\alpha} =$ 'r beta/alpha' : $\frac{d_{emp_2}}{d_{emp_1}}$:
#   $d_{hyp} =$ `r d_hyp[1]` \& $d_{true} =$ `r d_true` 
# `r mean(d_emp[1,which(LR[1,,1] >20 & Lhyp[1,,1] > 4 ),2]/d_hyp[1])`
# $d_{hyp} =$ `r d_hyp[2]` \& $d_{true} =$ `r d_true` 
# `r mean(d_emp[1,which(LR[2,,1] >20 & Lhyp[2,,1] > 4 ),2]/d_hyp[2])`

``` 
